Analyzing the Effects of the NBA COVID Bubble on the NBA Playoffs

### 2020-10-27 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

# III. Data Collection and Methods  
The 2020 bubble provides a new and exciting opportunity to study home-court
advantage for the NBA. Unlike college basketball, aside from a few 
exhibition/preseason games, the NBA always has a home and away team. So, for
the first time in NBA history the bubble allows NBA home and away performance
to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
removed all 4 possible game location factors impacting home-court advantage 
hypothesized by [@Carron2005]. The NBA bubble featured 8 seeding games then a
standard playoff format. The focus of this study was the play during the playoff
games since it followed the standard playoff format and can easily be compared
back to other playoffs. For this study, the 2020 playoffs were compared against
the three previous playoffs collectively. It is important to make sure that any
observed differences are truly from a home-court advantage. So, in order to control
for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019.
The reason being the faster pace play and more common use of the three-point shot 
in modern basketball. If we used data from say 10 years ago, or earlier, observed
differences may not be from effects of the NBA bubble, but rather from the effects
of drastic changes in the style of play between the seasons. However, basketball
evolves slow enough that we can reasonably assume 2017-2019 are at least very 
close in pace and playing style to 2020.

@@

(ref:cap-Figthree) Home shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)
```{r Figthree, echo=FALSE, fig.cap="(ref:cap-Figthree)"}
opar<-par(no.readonly = TRUE)
par(mfrow=c(2,3))

>>>>

# III. Data and Methods  
The 2020 bubble provides a new and exciting opportunity to study home-court
advantage for the NBA. Unlike college basketball, aside from a few 
exhibition/preseason games, the NBA always has a home and away team. So, for
the first time in NBA history the bubble allows NBA home and away performance
to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
removed all 4 possible game location factors impacting home-court advantage 
hypothesized by [@Carron2005]. The NBA bubble featured 8 seeding games then a
standard playoff format. The focus of this study was the play during the playoff
games since it followed the standard playoff format and can easily be compared
back to other playoffs. For this study, the 2020 playoffs were compared against
the three previous playoffs collectively. It is important to make sure that any
observed differences are truly from a home-court advantage. So, in order to control
for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019.
The reason being the faster pace play and more common use of the three-point shot 
in modern basketball. If we used data from say 10 years ago, or earlier, observed
differences may not be from effects of the NBA bubble, but rather from the effects
of drastic changes in the style of play between the seasons. However, basketball
evolves slow enough that we can reasonably assume 2017-2019 are at least very 
close in pace and playing style to 2020.

@@

```{r Figthree, echo=FALSE, fig.cap="Home shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)"}
opar<-par(no.readonly = TRUE)
par(mfrow=c(2,3))
